---
title: "Big N's"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Big N's}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, message = FALSE, warning = FALSE}
library(tfrmt)
library(tibble)
library(dplyr)
```

Displaying subject totals ("Big N's") at the top of a table is a very common and necessary part of building tables for clinical trials. {tfrmt} allows you to do this via the `big_n` argument, which allows you to use a `big_n_structure` object to specify which values in the ARD should be added to the columns and how they should be formatted. 


To dynamically add big N's to column labels, the information needs to be included in the ARD. As a simple example, let's take a small demographics table that just has Age and Sex for Placebo, Treatment and Total. 
Below is our ARD prior to incorporating the big Ns:
```{r}

data <- tibble(Group = rep(c("Age (y)", "Sex", "Age (y)", "Sex"), c(3, 3, 6,12)),
               Label = rep(c("n", "Mean (SD)", "Male","Female"), c(6, 6,6,6)),
               Column = rep(c("Placebo", "Treatment", "Total"), times = 8),
               Param = rep(c("n", "mean", "sd", "n", "pct", "n", "pct"),  c(6, 3, 3, 3,3,3,3)),
               Value = c(15,13,28,14,13,27,73.56, 74.231,71.84,9.347,7.234,8.293,8,7,15,8/14,7/13,15/27,6,6,12,6/14,6/13,12/27
               )
) %>%
  # Note because tfrmt only does rounding we will need to have the percents multiplied by 100
  mutate(Value = case_when(Param == "pct" ~ Value * 100,
                           TRUE ~ Value),
         ord1 = if_else(Group == "Age (y)", 1, 2),
         ord2 = if_else(Label == "n", 1, 2))
data
```


In this study we have 30 subjects per Arm, for a total of 60 subjects. To make the N's go to the right place we need to ensure the values in the `Column` column match the rest of the data. Then we can combine this information with teh data above to make the full ARD for this table.  
```{r}
big_ns <- tibble(Column = c("Placebo", "Treatment", "Total"),
                 Param = "bigN",
                 Value = c(30,30,60))


data <- bind_rows(data, big_ns)
```


Now we can make our `tfrmt` as normal, but this time we are going to add a `big_n_structure` to `big_n`. Here we will specify that values with the parameter of `"bigN"` should be in the column labels and they should be formatted as `"N = XX"` and on a seperate line.
```{r}
tfrmt(
  group = Group,
  label = Label,
  column = Column,
  value = Value,
  param = Param,
  sorting_cols = c(ord1, ord2),
  body_plan = body_plan(
    frmt_structure(group_val = ".default",
                   label_val = ".default",
                   frmt_combine("{n} {pct}",
                                n = frmt("X"),
                                pct = frmt("(xx.x%)", missing = " ")
                   )
    ),
    frmt_structure(group_val = "Age (y)", label_val = "Mean (SD)",
                   frmt_combine("{mean} ({sd})",
                                mean = frmt("XX.X"),
                                sd = frmt("x.xx")
                   )
    ),
    frmt_structure(group_val = ".default", label_val = "n", frmt("xx"))
  ),
  col_plan = col_plan(everything(), -starts_with("ord"), "Total"),
  row_grp_plan = row_grp_plan(
    row_grp_structure(group_val = ".default", element_block(post_space = " "))
  ),
  big_n = big_n_structure(param = "bigN", n_frmt = frmt("\nN = xx"))
) %>% 
  print_to_gt(data)
```

This will work no matter how complicated the table, but column names must always match in the ARD. For tables with spanning columns, column information up to the point where you want the big N to go. 
 Let's look at this table, which has 3 layers of column spanning
```{r}
  data <- tribble(
    ~group,     ~label,        ~span2,  ~span1,     ~my_col,    ~parm,   ~val,
    "g1", "rowlabel1",  "column cols", "cols 1,2", "col1"  ,  "value",    1,
    "g1", "rowlabel1",  "column cols", "cols 1,2", "col2"  ,  "value",    1,
    "g1", "rowlabel1",             NA,         NA, "mycol3",  "value",    1,
    "g1", "rowlabel1",  "column cols", "col 4"   , "col4"  ,  "value",    1,
    "g1", "rowlabel1",             NA,         NA, "mycol5",  "value",    1,
    "g1", "rowlabel2",  "column cols", "cols 1,2", "col1"  ,  "value",    2,
    "g1", "rowlabel2",  "column cols", "cols 1,2", "col2"  ,  "value",    2,
    "g1", "rowlabel2",             NA,        NA , "mycol3",  "value",    2,
    "g1", "rowlabel2",  "column cols", "col 4"   , "col4"  ,  "value",    2,
    "g1", "rowlabel2",             NA,         NA, "mycol5",  "value",    2,
    "g2", "rowlabel3",  "column cols", "cols 1,2", "col1"  ,  "value",    3,
    "g2", "rowlabel3",  "column cols", "cols 1,2", "col2"  ,  "value",    3,
    "g2", "rowlabel3",             NA,         NA, "mycol3",  "value",    3,
    "g2", "rowlabel3",  "column cols", "col 4"   , "col4"  ,  "value",    3,
    "g2", "rowlabel3",             NA,         NA, "mycol5",  "value",    3,
  )



spanning_tfrmt<- tfrmt(
    group = group,
    label = label,
    param = parm,
    values = val,
    column = c(span2, span1, my_col),
    body_plan = body_plan(
      frmt_structure(group_val = ".default", label_val = ".default", frmt("x"))
    ),
    col_plan = col_plan(
      group,
      label,
      starts_with("col")
    )
  )

print_to_gt(spanning_tfrmt, data)
```
 
 If we just wanted to put the big N on the highest "column cols" we would need to add a row to the ARD where `span2` equals `"column cols"` and all other column columns are missing. 
```{r}
with_big_n_data <- tribble(
  ~group,     ~label,        ~span2,  ~span1,     ~my_col,    ~parm,   ~val,
      NA,        NA,  "column cols",     NA,       NA  ,  "bigN",    18,
) %>% 
  bind_rows(data)

#Now we can add the big_n to the tfrmt from before 
n_span_tfrmt <- spanning_tfrmt %>% 
  tfrmt(big_n = big_n_structure(param = "bigN", n_frmt = frmt(" N = xx"))) 

print_to_gt(n_span_tfrmt, .data = with_big_n_data)
```
 
 If we also wanted big N's on `col 1,2`, `col4`, and `mycol3`. So we will need Add 3 more rows to our ARD. One that goes down to teh `span1` level of specificity and two that go all the way down to `my_col` level of specificity. (Note: because we are keeping the `parm` values the same we can use the same tfrmt)
```{r}
with_more_big_n_data <- tribble(
  ~group,     ~label,        ~span2,  ~span1,     ~my_col,    ~parm,   ~val,
    NA,         NA,  "column cols",  "cols 1,2",   NA  ,      "bigN",    12,
    NA,         NA,  "column cols",  "col 4"   ,   "col4"  ,  "bigN",    6,
    NA,         NA,             NA,          NA,   "mycol3",  "bigN",    6
) %>% 
  bind_rows(with_big_n_data)

print_to_gt(n_span_tfrmt, .data = with_more_big_n_data)
```
 
As a final note, if column names get changed in the `col_plan`, it shouldn't effect the ARD. The column names for the big N's should match the ARD column names not any future names. But, if names do get updated the big N's will be uneffected. 
```{r}
n_span_tfrmt %>% 
  tfrmt(col_plan = col_plan(
      group,
      label,
      starts_with("col"),
      new_col_3 = mycol3,
      -mycol5
    )) %>% 
  print_to_gt(.data = with_more_big_n_data)

```


`big_n_structure`s provide a straightforward way of adding big N's from the ARD into the column labels. The functionality allows for big N's at any level of column spanning depending on how it is parameterized in the ARD. 
