---
title: "Big N's"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Big N's}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, message = FALSE, warning = FALSE}
library(tfrmt)
library(tibble)
library(dplyr)
library(tidyr)
```

Displaying subject totals ("Big N's") at the top of a table is a very common and necessary part of building tables for clinical trials. {tfrmt} allows you to do this via the `big_n` argument, which allows you to use a `big_n_structure` object to specify which values in the ARD should be added to the columns and how they should be formatted. 

## Overall Big Ns

To dynamically add big N's to column labels, the information needs to be included in the ARD. As a simple example, let's take a small demographics table that just has Age and Sex for Placebo, Treatment and Total. 
Below is our ARD prior to incorporating the big Ns:
```{r}
data <- tibble(
  Group = rep(c("Age (y)", "Sex", "Age (y)", "Sex"), c(3, 3, 6, 12)),
  Label = rep(c("n", "Mean (SD)", "Male", "Female"), c(6, 6, 6, 6)),
  Column = rep(c("Placebo", "Treatment", "Total"), times = 8),
  Param = rep(c("n", "mean", "sd", "n", "pct", "n", "pct"), c(6, 3, 3, 3, 3, 3, 3)),
  Value = c(
    15, 13, 28, 14, 13, 27, 73.56, 74.231, 71.84, 9.347, 7.234, 8.293,
    8, 7, 15, 8 / 14, 7 / 13, 15 / 27, 6, 6, 12, 6 / 14, 6 / 13, 12 / 27
  )
) %>%
  # Note because tfrmt only does rounding we will need to have the percents multiplied by 100
  mutate(
    Value = case_when(
      Param == "pct" ~ Value * 100,
      TRUE ~ Value
    ),
    ord1 = if_else(Group == "Age (y)", 1, 2),
    ord2 = if_else(Label == "n", 1, 2)
  )
data
```


In this study we have 30 subjects per treatment arm for a total of 60 subjects. Because the ARD is 1 row per unique value, our Ns will contribute 3 rows to the ARD. To make the N's go to the right place, we need to supply information about the `column` column values. The `column` column values should map to those already available in the ARD. Then we can combine this information with the data above to make the full ARD for this table.  
```{r}
big_ns <- tibble(
  Column = c("Placebo", "Treatment", "Total"),
  Param = "bigN",
  Value = c(30, 30, 60)
)


data <- bind_rows(data, big_ns)
```


Now we can make our `tfrmt` as normal, but this time we are going to add a `big_n_structure` to `big_n`. Here we will specify that values with the parameter of `"bigN"` should be appended to the column labels and they should be formatted as `"N = XX"` and on a separate line.
```{r}
tfrmt(
  group = Group,
  label = Label,
  column = Column,
  value = Value,
  param = Param,
  sorting_cols = c(ord1, ord2),
  body_plan = body_plan(
    frmt_structure(
      group_val = ".default",
      label_val = ".default",
      frmt_combine("{n} {pct}",
        n = frmt("X"),
        pct = frmt("(xx.x%)", missing = " ")
      )
    ),
    frmt_structure(
      group_val = "Age (y)", label_val = "Mean (SD)",
      frmt_combine("{mean} ({sd})",
        mean = frmt("XX.X"),
        sd = frmt("x.xx")
      )
    ),
    frmt_structure(group_val = ".default", label_val = "n", frmt("xx"))
  ),
  col_plan = col_plan(everything(), -starts_with("ord"), "Total"),
  row_grp_plan = row_grp_plan(
    row_grp_structure(group_val = ".default", element_block(post_space = " "))
  ),
  big_n = big_n_structure(param_val = "bigN", n_frmt = frmt("\nN = xx"))
) %>%
  print_to_gt(data)
```

No matter how complicated the table, this feature will work as long as the ARD is set up appropriately with the correct `column` column values on the big N rows. For tables with spanning columns (i.e. multiple `column` columns), the big N will be appended to the lowest level `column` column value that is supplied (based on the order of the column variables supplied to the `tfrmt`).
 Let's look at this table, which has 3 layers of column spanning:
```{r}
data <- tribble(
  ~group, ~label, ~span2, ~span1, ~my_col, ~parm, ~val,
  "g1", "rowlabel1", "column cols", "cols 1,2", "col1", "value", 1,
  "g1", "rowlabel1", "column cols", "cols 1,2", "col2", "value", 1,
  "g1", "rowlabel1", NA, NA, "mycol3", "value", 1,
  "g1", "rowlabel1", "column cols", "col 4", "col4", "value", 1,
  "g1", "rowlabel1", NA, NA, "mycol5", "value", 1,
  "g1", "rowlabel2", "column cols", "cols 1,2", "col1", "value", 2,
  "g1", "rowlabel2", "column cols", "cols 1,2", "col2", "value", 2,
  "g1", "rowlabel2", NA, NA, "mycol3", "value", 2,
  "g1", "rowlabel2", "column cols", "col 4", "col4", "value", 2,
  "g1", "rowlabel2", NA, NA, "mycol5", "value", 2,
  "g2", "rowlabel3", "column cols", "cols 1,2", "col1", "value", 3,
  "g2", "rowlabel3", "column cols", "cols 1,2", "col2", "value", 3,
  "g2", "rowlabel3", NA, NA, "mycol3", "value", 3,
  "g2", "rowlabel3", "column cols", "col 4", "col4", "value", 3,
  "g2", "rowlabel3", NA, NA, "mycol5", "value", 3,
)



spanning_tfrmt <- tfrmt(
  group = group,
  label = label,
  param = parm,
  value = val,
  column = c(span2, span1, my_col),
  body_plan = body_plan(
    frmt_structure(group_val = ".default", label_val = ".default", frmt("x"))
  ),
  col_plan = col_plan(
    group,
    label,
    starts_with("col")
  )
)

print_to_gt(spanning_tfrmt, data)
```
 
 If we just want to put the big N on the highest "column cols", we need to add a row to the ARD where `span2` equals `"column cols"` and all other `column` columns are missing. 
```{r}
with_big_n_data <- tribble(
  ~group, ~label, ~span2, ~span1, ~my_col, ~parm, ~val,
  NA, NA, "column cols", NA, NA, "bigN", 18,
) %>%
  bind_rows(data)

# Now we can add the big_n to the tfrmt from before
n_span_tfrmt <- spanning_tfrmt %>%
  tfrmt(big_n = big_n_structure(param_val = "bigN", n_frmt = frmt("\nN = xx")))

print_to_gt(n_span_tfrmt, .data = with_big_n_data)
```
 
 If we also want big N's on `col 1,2`, `col4`, and `mycol3`, we will need Add 3 more rows to our ARD. One that goes down to the `span1` level of specificity and two that go all the way down to `my_col` level of specificity. (Note: because we are keeping the `parm` values the same we can use the same tfrmt)
```{r}
with_more_big_n_data <- tribble(
  ~group, ~label, ~span2, ~span1, ~my_col, ~parm, ~val,
  NA, NA, "column cols", "cols 1,2", NA, "bigN", 12,
  NA, NA, "column cols", "col 4", "col4", "bigN", 6,
  NA, NA, NA, NA, "mycol3", "bigN", 6
) %>%
  bind_rows(with_big_n_data)

print_to_gt(n_span_tfrmt, .data = with_more_big_n_data)
```
 
## Page-Level Big Ns

The `page_plan` allows users to split up large tables in many ways. Let's suppose our `page_plan` splits our table so that each subtable represents a different subpopulation defined by the `group` variable in the ARD. 

In this case, our page_plan looks something like this:
```r
page_structure(group_val = ".default")
```

Because our subtables contain different populations, we want the big Ns displayed in the headers to be table-specific. To achieve this, we can supply multiple sets of big Ns in the ARD. We just need to specify the specific values of the `group` parameter that each big N belongs to. This functionality can be enabled via the `by_page` argument in `big_n_structure`.

For example, let's take the following data, where we have calculated big Ns for each level of the grouping variable `grp`. NOTE: the bigN rows should be in the same order as the rest of the data (i.e. A, B rather than B, A for the `grp` variable).
```{r}
dat <- crossing(
  grp = c("A", "B"),
  lbl = c("a", "b"),
  col = c("Placebo", "Treatment"),
  param = "mean"
) %>%
  mutate(val = c(1.254, 3.483, 5.123, 4.239, 4.364, 8.435, 7.645, 2.312))

big_ns <- tibble(
  col = c("Placebo", "Placebo", "Treatment", "Treatment"),
  grp = c("A", "B", "A", "B"),
  param = "bigN",
  val = c(34, 36, 42, 39)
)


dat <- bind_rows(dat, big_ns)
tail(dat)
``` 

Now that we have all big Ns represented in the ARD, we can make sure of the `by_page` argument in the `big_n_structure` to ensure the various big Ns are assigned to the relevant pages defined by the `page_plan`.
```{r}
gts <- tfrmt(
  group = grp,
  label = lbl,
  column = col,
  param = param,
  value = val,
  body_plan = body_plan(
    frmt_structure(group_val = ".default", label_val = ".default", frmt("x.x"))
  ),
  page_plan = page_plan(
    page_structure(group_val = ".default")
  ),
  big_n = big_n_structure(param_val = "bigN", n_frmt = frmt("\nN = xx"), by_page = TRUE)
) %>%
  print_to_gt(dat)
```

```{r}
gts %>% gt::grp_pull(1)
```
```{r}
gts %>% gt::grp_pull(2)
``` 


 
## Final note

As a final note, column names for the big N's should match the ARD column names, rather than any future names. Regardless of any renaming that happens in the `col_plan`, the column name for the big N's should be consistent with column values in the ARD. The big N's will still be added if names do get updated in the `col_plan`.  
```{r}
n_span_tfrmt %>%
  tfrmt(col_plan = col_plan(
    group,
    label,
    starts_with("col"),
    new_col_3 = mycol3,
    -mycol5
  )) %>%
  print_to_gt(.data = with_more_big_n_data)
```


`big_n_structure`s provide a straightforward way of adding big N's from the ARD into the column labels. The functionality allows for big N's at any level of column spanning depending on how it is parameterized in the ARD. 
