---
title: "Body Plan"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Body Plan}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, echo = FALSE}
library(tfrmt)
```

The body plan covers the formatting of numeric cell values. Within the body plan, format structures (`frmt_structure`) may be layered to specify formatting for different parameters. Formatting can be applied to specific group and/or label values, which we will cover in a bit. First, we will apply formats to all relevant values in the table, regardless of group or label value.

# Formatting functions

The following formatting functions may be passed to `frmt_structure`:

## Format (`frmt`)

The `frmt` object is used to specify how a value should appear in a table: its number of digits, any white space padding, or extra characters, such as percents or parentheses. In pharma, it is a common convention to use x's to represent data values in mock tables. As such, the `frmt` complies with this familiar convention. For example:

- `frmt(expression = "xxx.x%")` displays 3.829765 as "  3.8%" (note the leading whitespace!)
- `frmt(expression = "xxx.x%", missing = "Missing")` displays NA as "Missing"
- `frmt(expression = "xxx.x", scientific = " x10^x")` displays 1234.5678 as "  1.2 x10^ 3"

## Format when (`frmt_when`)

`frmt_when` is a way to format cells conditional on their values. `frmt_when` is similar to dplyr::case_when or if/else statements. For each condition, the left hand side is an expression as a string and the right hand side is the formatting to be applied. For example, the following code allows users to conditionally format the upper and lower bounds for p-values:

```
frmt_when(
    ">0.99" ~ ">0.99",
    "<0.001" ~ "<0.001",
    TRUE ~ frmt("x.xxx")
)
```

This will cause the values 0.0347, 1.00, and 0.000001 to appear as "0.035", ">0.99", and "<0.001".

## Format combine (`frmt_combine`)

`frmt_combine` is used to combine multiple values into a single cell. Within a `frmt_combine`, the values are referenced according to the values of the `param` variable. Glue package syntax is leveraged to help the user easily define how they want the combined values to appear. Further, the individual data values may be formatted using `frmt` or `frmt_when`. Some examples:

- `frmt_combine("({lower}, {upper})", lower = frmt("xx.x"), upper = frmt("xx.x"))` will combine lower and upper bounds of a confidence interval into: "(xx.x, xx.x)".
 
- `frmt_combine("{n} ({pct})", n = frmt("xxx"), pct = frmt("xx.x", missing = ""))` will print "n (%)" if the percentage is non-missing, but just "n" otherwise. 

- `frmt_combine("{n} ({pct})", n = frmt("xxx"), pct = frmt("xx.x", missing = ""), missing = "")` will print "n (%)" if the percentage is non-missing, "n" if the percentage is missing but n is non-missing, and "" if both n and pct are missing. 

# Location options

There are a variety of ways for the user to specify formatting locations: group variable values and/or label values, or by param value. Let's take a look at the various methods below, ranging from least specific to most specific. 

An important details is that passing a value of ".default" to group or label means that the format will be applied to ALL values within the group(s) or label variables. 

## Table-wide

A user can specify formatting for the entire table as such:

`frmt(group = ".default", label = ".default", frmt("xx.x"))`

## Param-specific

For formatting specific values of the param variable, `frmt` objects must be named according to the param. The following code will format all values where the param value equals "pval", regardless of group, label, or column value.

`frmt(group = ".default", label = ".default", pval = frmt("x.xxx"))`

Again, param names are required for using `frmt_combine`.

## Label-specific

For formatting specific values of the label variable, these values are passed to the `label` argument. The following code will format all values where the label variable equals "mylab1" or "mylab2", regardless of group, param, or column value.

`frmt(group = ".default", label = c("mylab1","mylab2"), frmt("x.xxx"))`

## Group-specific

For formatting specific values of the group variable, these values are passed to the `group` argument. Note that the user may have multiple group variables. In the case of multiple group variables, the specifications are to be passed as a named list (named according to the group variables). The following code will format all values where group1="A" and group2="a" or "b", regardless of label, param, or column value:

`frmt(group = list(group1 = "A", group2 = "B"), label = ".default", frmt("x.xxx"))`

## Column-specific
In the case where formatting needs to differ between columns within a given group/label/param level, the user will need to incorporate column-level information into the `param` value. For example, if we have two columns, `col1` and `col2`, but we want the mean for `col1` to be rounded differently than `col2`, creating our `param` column as such will allow us to differentiate the mean values in the formatting: `param = c("col1_mean", "col2_mean")`. 

## Multiple specifications

A user can be as specific as they'd like, by providing any of these in combination. For example, the following code will format values where group1="A", group2="A", the label variable equals "mylab" and the param variable equals "mean":

`frmt(group = list(group1 = "A", group2 = "B"), label = "mylab", mean = frmt("x.xxx"))`

**NOTE:** the `tfrmt_sigdig()` function provides the ability to have group and label value-specific formatting. 

# Layering format structures

The user may provide as many `frmt_structure` objects as needed. It is important to note that if there are overlapping `frmt_structure` objects, the last one specified will supercede the previous. In our CDISC demographics table example, the following `frmt_structures` are needed to cover all values:

```
body_plan(
    frmt_structure(group_val = ".default", label_val = ".default",
                   frmt_combine("{n} {pct}",
                                n = frmt("XXX"),
                                pct = frmt_when("==100" ~ "",
                                                "==0" ~ "",
                                                TRUE ~ frmt("(XX.X %)"))
                   )
    ),
    frmt_structure(group_val = ".default", label_val = "n", frmt("xxx")),
    frmt_structure(group_val = ".default", label_val = c("Mean", "Median", "Min","Max"), frmt("xxx.x")),
    frmt_structure(group_val = ".default", label_val = "SD", frmt("xxx.xx")),
    frmt_structure(group_val = ".default", label_val = ".default", p = frmt_when(">0.99" ~ ">0.99",
                                                                                 "<0.001" ~ "<0.001",
                                                                                 TRUE ~ frmt("x.xxx", missing = "")))
)
```
