---
title: "Column Plan"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Column Plan}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, message = FALSE, warning = FALSE}
library(tfrmt)
library(tibble)
library(dplyr)
```

The purpose of the column plan (`col_plan`) is to allow the user to specify how columns in the final table should appear in regards to their labels and ordering. Similar to the column styling plan (`col_style_plan`), `col_plan` refers to the final columns of the table, i.e., the values found in the `column` variable(s) from the input dataset. The `col_plan` is like an advanced version of dplyr's `select` function; just like `select`, the user can add, drop, order, and/or rename columns all in one step. What sets the `col_plan` apart, however, is the ability to also specify column spanners and labels in a hierarchical fashion, as well as preferentially selecting the _last_ instance a column is selected. For column spanning, multiple column columns need to exist in the input dataset. 


Let's look at the workflow in more detail with the following dataset which includes a single `column` variable:

```{r}
dat <- tribble(
  ~group,     ~label,  ~my_col,    ~parm, ~val,
  "g1", "rowlabel1",  "col1"  ,  "value",    1,
  "g1", "rowlabel1",  "col2"  ,  "value",    1,
  "g1", "rowlabel1",  "mycol3",  "value",    1,
  "g1", "rowlabel1",  "col4"  ,  "value",    1,
  "g1", "rowlabel1",  "mycol5",  "value",    1,
  "g1", "rowlabel2",  "col1"  ,  "value",    2,
  "g1", "rowlabel2",  "col2"  ,  "value",    2,
  "g1", "rowlabel2",  "mycol3",  "value",    2,
  "g1", "rowlabel2",  "col4"  ,  "value",    2,
  "g1", "rowlabel2",  "mycol5",  "value",    2,
  "g2", "rowlabel3",  "col1"  ,  "value",    3,
  "g2", "rowlabel3",  "col2"  ,  "value",    3,
  "g2", "rowlabel3",  "mycol3",  "value",    3,
  "g2", "rowlabel3",  "col4"  ,  "value",    3,
  "g2", "rowlabel3",  "mycol5",  "value",    3)

```


# Simple Column Selection

In the case of a single column variable with no column spanners, `col_plan` behaves similarly to `dplyr::select`. 

Let's say we want to remove one of the columns:
```{r}
tfrmt(
  group = group,
  label = label,
  param = parm,
  value = val,
  column = my_col,
  body_plan = body_plan(
    frmt_structure(group_val = ".default", label_val = ".default", frmt("x"))
  ),
  col_plan = col_plan(
    -mycol5            
  )
) %>% 
  print_to_gt(dat)
```

Just like `dplyr::select`, we can also reorder, rename, and remove columns, using `tidyselect` syntax if we'd like:

```{r}
tfrmt(
  group = group,
  label = label,
  param = parm,
  value = val,
  column = my_col,
  body_plan = body_plan(
    frmt_structure(group_val = ".default", label_val = ".default", frmt("x"))
  ),
  col_plan = col_plan(
    group, 
    label, 
    starts_with("col"),
    new_col_3 = mycol3,  
    -mycol5            
  )
) %>% 
  print_to_gt(dat)
```

Unlike `dplyr::select` though, col_plan respects the _last_ time a column is defined in the plan. This allows for reordering/moving a column to the end very easily.

```{r}
tfrmt(
  group = group,
  label = label,
  param = parm,
  value = val,
  column = my_col,
  body_plan = body_plan(
    frmt_structure(group_val = ".default", label_val = ".default", frmt("x"))
  ),
  col_plan = col_plan(
    group, 
    label, 
    starts_with("col"),
    everything(),
    col1         
  )
) %>% 
  print_to_gt(dat)
```


# Editing and Moving Column Spanners

Multiple `column` variables are used to form the hierarchy of column spanners and column labels, which is driven by the order of the `column` variables. The first variable specified represents the highest level spanner, while the last variable specified represents the lowest level column label.

Let's consider an example dataset with multiple `column` variables. This dataset has three tiers, `span2` with one unique value, `span1` with two unique values and `my_col` which has 4 unique values. The `NA`'s indicate there should be no spanning over those values. 

```{r}
dat <- tribble(
  ~group,     ~label,        ~span2,  ~span1,     ~my_col,    ~parm,   ~val,
  "g1", "rowlabel1",  "column cols", "cols 1,2", "col1"  ,  "value",    1,
  "g1", "rowlabel1",  "column cols", "cols 1,2", "col2"  ,  "value",    1,
  "g1", "rowlabel1",             NA,         NA, "mycol3",  "value",    1,
  "g1", "rowlabel1",  "column cols", "col 4"   , "col4"  ,  "value",    1,
  "g1", "rowlabel1",             NA,         NA, "mycol5",  "value",    1,
  "g1", "rowlabel2",  "column cols", "cols 1,2", "col1"  ,  "value",    2,
  "g1", "rowlabel2",  "column cols", "cols 1,2", "col2"  ,  "value",    2,
  "g1", "rowlabel2",             NA,        NA , "mycol3",  "value",    2,
  "g1", "rowlabel2",  "column cols", "col 4"   , "col4"  ,  "value",    2,
  "g1", "rowlabel2",             NA,         NA, "mycol5",  "value",    2,
  "g2", "rowlabel3",  "column cols", "cols 1,2", "col1"  ,  "value",    3,
  "g2", "rowlabel3",  "column cols", "cols 1,2", "col2"  ,  "value",    3,
  "g2", "rowlabel3",             NA,         NA, "mycol3",  "value",    3,
  "g2", "rowlabel3",  "column cols", "col 4"   , "col4"  ,  "value",    3,
  "g2", "rowlabel3",             NA,         NA, "mycol5",  "value",    3,
)
```

Similar to the case of 1 `column` variable, the user may remove and rename values for the lowest level `column` variable within col_plan. To edit or move the columns based on the the column spanners, span_structure must be used. Here we want to bring all the columns that start with "col" together, as well as rename `mycol3` and drop `mycol5`. 

```{r}
tfrmt(
  group = group,
  label = label,
  param = parm,
  value = val,
  column = c(span2, span1, my_col),
  body_plan = body_plan(
    frmt_structure(group_val = ".default", label_val = ".default", frmt("x"))
  ),
  col_plan = col_plan(
    group,
    label,
    starts_with("col"),
    new_col_3 = mycol3,
    -mycol5
  )
) %>% 
  print_to_gt(dat)
```


Let's suppose we want to move the "col 4" column to the beginning and also reorder "col1" and "col2". To achieve this, we can use `span_structure`. `span_structure` allows you to specify value(s) for a given column column, to specify the order. So can select `"col 4"`  from `span1`. Then use another `span_strcture` to select `"col2"` and then `"col1"` from `my_col` when `span1` equals `"col 1,2"`,and finally all the other columns. 

```{r}
tfrmt(
  group = group,
  label = label,
  param = parm,
  value = val,
  column = c(span2, span1, my_col),
  body_plan = body_plan(
    frmt_structure(group_val = ".default", label_val = ".default", frmt("x"))
  ),
  col_plan = col_plan(
    group, 
    label, 
    span_structure(span1 = c("col 4")),
    span_structure(span1 = c("cols 1,2"), my_col = c("col2", "col1")),
    everything(),
    new_col_3 = mycol3,  
    -mycol5            
  )
) %>% 
  print_to_gt(dat)
```


Reordering with multiple `column` variables is simplest when the lowest level column variable contains unique values (i.e., there is 1 value per column in the final table), like in the example above. But, that isn't always the case take the following example:

```{r}
dat <- tribble(
  ~group,     ~label,     ~span ,  ~my_col,    ~parm,   ~val,
  "g1",      "stats",  "Placebo" , "sd"     ,  "sd"  ,    1.435,
  "g1",      "stats",  "Placebo" , "mean"   ,  "mean",    2.843,
  "g1",      "stats",  "Treatment", "mean"   ,  "mean",    1.234,
  "g1",      "stats",  "Treatment", "sd"     ,  "sd"  ,    2.123,
)
```

For this table we want to ensure `"Treatment"` is always before `"Placebo"` and that `"mean"` comes before `"sd"`. To do that we can make one `span_structure` that specifies both the order of the `span` level and the order of the `my_col`  level. 

```{r}
tfrmt(
  group = group,
  label = label,
  param = parm,
  value = val,
  column = c(span, my_col),
  body_plan = body_plan(
    frmt_structure(group_val = ".default", label_val = ".default", frmt("x.xx"))
  ),
  col_plan = col_plan(
    group, 
    label, 
    span_structure(span = c("Treatment", "Placebo"),
                   my_col = c("mean", "sd"))
  )
) %>% 
  print_to_gt(dat)
```

For more examples of `col_plan`s see the examples or unusual table vignettes. 





