---
title: "Column Plan"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Column Plan}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, message = FALSE, warning = FALSE}
library(tfrmt)
library(tibble)
library(dplyr)
```

The purpose of the column plan (`col_plan`) is to allow the user to specify how columns in the final table should appear in regards to their labels and ordering. Similar to the column styling plan (`col_style_plan`), `col_plan` refers to the final columns of the table, i.e., the values found in the `column` variable(s) from the input dataset. The `col_plan` is like an advanced version of dplyr's `select` function; just like `select`, the user can add, drop, order, and/or rename columns all in one step. What sets the `col_plan` apart, however, is the ability to also specify column spanners and labels in a hierarchical fashion, as well as preferentially selecting the _last_ instance a column is selected. For column spanning, multiple column columns need to exist in the input dataset. 

Let's look at the workflow in more detail with the following dataset which includes a single `column` variable:

```{r}
dat <- tribble(
  ~group, ~label, ~my_col, ~parm, ~val,
  "g1", "rowlabel1", "col1", "value", 1,
  "g1", "rowlabel1", "col2", "value", 1,
  "g1", "rowlabel1", "mycol3", "value", 1,
  "g1", "rowlabel1", "col4", "value", 1,
  "g1", "rowlabel1", "mycol5", "value", 1,
  "g1", "rowlabel2", "col1", "value", 2,
  "g1", "rowlabel2", "col2", "value", 2,
  "g1", "rowlabel2", "mycol3", "value", 2,
  "g1", "rowlabel2", "col4", "value", 2,
  "g1", "rowlabel2", "mycol5", "value", 2,
  "g2", "rowlabel3", "col1", "value", 3,
  "g2", "rowlabel3", "col2", "value", 3,
  "g2", "rowlabel3", "mycol3", "value", 3,
  "g2", "rowlabel3", "col4", "value", 3,
  "g2", "rowlabel3", "mycol5", "value", 3
)
```

This is what this data looks like formatted as a basic table.

```{r}
tfrmt(
  group = group,
  label = label,
  param = parm,
  value = val,
  column = my_col,
  body_plan = body_plan(
    frmt_structure(group_val = ".default", label_val = ".default", frmt("x"))
  )
) %>%
  print_to_gt(dat)
```

# Simple Column Selection

In the case of a single column variable with no column spanners, `col_plan` behaves similarly to `dplyr::select`. 

If we want to remove one of the columns we specify within a `col_plan`:

```{r}
tfrmt(
  group = group,
  label = label,
  param = parm,
  value = val,
  column = my_col,
  body_plan = body_plan(
    frmt_structure(group_val = ".default", label_val = ".default", frmt("x"))
  ),
  col_plan = col_plan(
    -mycol5
  )
) %>%
  print_to_gt(dat)
```

Just like `dplyr::select`, we can also reorder, rename, and remove columns, using `tidyselect` syntax if we'd like:

```{r}
tfrmt(
  group = group,
  label = label,
  param = parm,
  value = val,
  column = my_col,
  body_plan = body_plan(
    frmt_structure(group_val = ".default", label_val = ".default", frmt("x"))
  ),
  col_plan = col_plan(
    # reordering
    group,
    label,
    starts_with("col"),
    # renaming
    new_col_3 = mycol3,
    # removing
    -mycol5
  )
) %>%
  print_to_gt(dat)
```

Unlike `dplyr::select` though, col_plan respects the _last_ time a column is defined in the plan. This allows for reordering/moving a column to the end very easily.

```{r}
tfrmt(
  group = group,
  label = label,
  param = parm,
  value = val,
  column = my_col,
  body_plan = body_plan(
    frmt_structure(group_val = ".default", label_val = ".default", frmt("x"))
  ),
  col_plan = col_plan(
    group,
    label,
    starts_with("col"),
    everything(),
    col1 # moved to end
  )
) %>%
  print_to_gt(dat)
```

# Naming the Group-Label Header

It is also possible to provide a combined header for the group and label columns - this is termed the "stub" in {gt}. To achieve this, rename the `group` variable in the column plan, just as you would with other columns. If multiple `group` variables exist, any of them can be renamed. If more than one is renamed, {tfrmt} will use the highest level `group` name available. 

```{r}
tfrmt(
  group = group,
  label = label,
  param = parm,
  value = val,
  column = my_col,
  body_plan = body_plan(
    frmt_structure(group_val = ".default", label_val = ".default", frmt("x"))
  ),
  col_plan = col_plan(
    my_grp = group, # rename group
    label,
    starts_with("col"),
    everything(),
    col1
  )
) %>%
  print_to_gt(dat)
```

Note that it is not currently possible to provide individual column headers if there are multiple group/label columns. 

# Editing and Moving Column Spanners

Multiple `column` variables are used to form the hierarchy of column spanners and column labels, which is driven by the order of the `column` variables. The first variable specified represents the highest level spanner, while the last variable specified represents the lowest level column label.

Let's consider an example dataset with multiple `column` variables. This dataset has three tiers, `span2` with one unique value, `span1` with two unique values and `my_col` which has 4 unique values. The `NA`'s indicate there should be no spanning over those values. 

```{r}
dat <- tribble(
  ~group, ~label, ~span2, ~span1, ~my_col, ~parm, ~val,
  "g1", "rowlabel1", "column cols", "cols 1,2", "col1", "value", 1,
  "g1", "rowlabel1", "column cols", "cols 1,2", "col2", "value", 1,
  "g1", "rowlabel1", NA, NA, "mycol3", "value", 1,
  "g1", "rowlabel1", "column cols", "col 4", "col4", "value", 1,
  "g1", "rowlabel1", NA, NA, "mycol5", "value", 1,
  "g1", "rowlabel2", "column cols", "cols 1,2", "col1", "value", 2,
  "g1", "rowlabel2", "column cols", "cols 1,2", "col2", "value", 2,
  "g1", "rowlabel2", NA, NA, "mycol3", "value", 2,
  "g1", "rowlabel2", "column cols", "col 4", "col4", "value", 2,
  "g1", "rowlabel2", NA, NA, "mycol5", "value", 2,
  "g2", "rowlabel3", "column cols", "cols 1,2", "col1", "value", 3,
  "g2", "rowlabel3", "column cols", "cols 1,2", "col2", "value", 3,
  "g2", "rowlabel3", NA, NA, "mycol3", "value", 3,
  "g2", "rowlabel3", "column cols", "col 4", "col4", "value", 3,
  "g2", "rowlabel3", NA, NA, "mycol5", "value", 3,
)
```

Similar to the case of 1 `column` variable, the user may remove and rename values for the lowest level `column` variable within col_plan. To edit or move the columns based on the the column spanners, `span_structure` must be used. Here we want to bring all the columns that start with "col" together, as well as rename `mycol3` and drop `mycol5`. 

```{r}
tfrmt(
  group = group,
  label = label,
  param = parm,
  value = val,
  # specify spanner columns and unique column from data
  column = c(span2, span1, my_col),
  body_plan = body_plan(
    frmt_structure(group_val = ".default", label_val = ".default", frmt("x"))
  ),
  col_plan = col_plan(
    group,
    label,
    starts_with("col"),
    new_col_3 = mycol3,
    -mycol5
  )
) %>%
  print_to_gt(dat)
```

Renaming column spanners can also be specified using `span_structure`. We can rename multiple spanners _within the same level_ in the same `span_structure` - here `"cols 1,2"` becomes `"first cols"` and `"col 4"` becomes `"just col4"` for the `span1` level. But when renaming multiple spanners _in different levels_ we need to create a new `span_structure` for each level - so we write `"column cols"` becomes `"most columns"` for the `span2` level in a separate `span_structure`.

```{r}
tfrmt(
  group = group,
  label = label,
  param = parm,
  value = val,
  column = c(span2, span1, my_col),
  body_plan = body_plan(
    frmt_structure(group_val = ".default", label_val = ".default", frmt("x"))
  ),
  col_plan = col_plan(
    # rename column spanner in same level
    span_structure(span1 = c(
      "first cols" = "cols 1,2",
      "just col4" = "col 4"
    )),
    # rename column spanner in different level
    span_structure(span2 = c("most columns" = "column cols")),
    group,
    label,
    starts_with("col"),
    new_col_3 = mycol3,
    -mycol5
  )
) %>%
  print_to_gt(dat)
```

Let's suppose we want to move the "col 4" column to the beginning and also reorder "col1" and "col2". To achieve this, we can use `span_structure`. `span_structure` allows you to specify value(s) for a given column column, to specify the order. So can select `"col 4"`  from `span1`. Then use another `span_strcture` to select `"col2"` and then `"col1"` from `my_col` when `span1` equals `"col 1,2"`,and finally all the other columns. 

```{r}
tfrmt(
  group = group,
  label = label,
  param = parm,
  value = val,
  column = c(span2, span1, my_col),
  body_plan = body_plan(
    frmt_structure(group_val = ".default", label_val = ".default", frmt("x"))
  ),
  col_plan = col_plan(
    group,
    label,
    # reordering spanners
    span_structure(span1 = c("col 4")),
    span_structure(span1 = c("cols 1,2"), my_col = c("col2", "col1")),
    everything(),
    new_col_3 = mycol3,
    -mycol5
  )
) %>%
  print_to_gt(dat)
```

Reordering with multiple `column` variables is simplest when the lowest level column variable contains unique values (i.e., there is 1 value per column in the final table), like in the example above. But, that isn't always the case take the following example:

```{r}
dat <- tribble(
  ~group, ~label, ~span, ~my_col, ~parm, ~val,
  "g1", "stats", "Placebo", "sd", "sd", 1.435,
  "g1", "stats", "Placebo", "mean", "mean", 2.843,
  "g1", "stats", "Treatment", "mean", "mean", 1.234,
  "g1", "stats", "Treatment", "sd", "sd", 2.123,
)
```

For this table we want to ensure `"Treatment"` is always before `"Placebo"` and that `"mean"` comes before `"sd"`. To do that we can make one `span_structure` that specifies both the order of the `span` level and the order of the `my_col`  level. 

```{r}
tfrmt(
  group = group,
  label = label,
  param = parm,
  value = val,
  column = c(span, my_col),
  body_plan = body_plan(
    frmt_structure(group_val = ".default", label_val = ".default", frmt("x.xx"))
  ),
  col_plan = col_plan(
    group,
    label,
    span_structure(
      span = c("Treatment", "Placebo"),
      my_col = c("mean", "sd")
    )
  )
) %>%
  print_to_gt(dat)
```

For more examples of `col_plan` see the examples or unusual table vignettes. 


