---
title: "Column Style Plan"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Column Style Plan}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, message = FALSE, warning = FALSE}
library(tfrmt)
```

tfrmt offers the ability to align value and set column widths. Columns can be aligned in the following ways:

* Character alignment
  -   Left-align
  -   Right-align
  -   Align on one or more single characters (e.g., decimal alignment)

* Positional alignment
  -   Align by any number of positions across the formatted values

Column widths are set as either the number of pixels wide they are, or the percentage of the total table with.

Alignment and width is specified via the `col_style_plan`, which accepts a series of `col_style_structure`s as inputs.

Alignment is achieved by padding the values with character spaces; this padding ensures that alignment is robust and agnostic to the output format. Width is done during post-processing and creation of the table output.

## Style Structure

`col_style_structure` has four arguments: the columns in the final table to be aligned, the alignment desired, the alignment type, and the widths to apply to those columns. Each `col_style_structure` requires at least one of alignment and width to be defined. Optionally both can be assigned too.

The column values that are specified are to match the contents of the `column` variable that is specified in the `tfrmt` object. When multiple values are provided to `column`, the last value is used as the reference column.

When the `col_style_plan()` is applied to the data, the most recent alignment and width is applied to the data, so be careful when adding new `col_style_structure`'s, as the output may change more than anticipated if the new element is not specific enough.

## Alignment

### Alignment Examples - Character

Let's take a look at how alignment can be applied to the example below, which contains a variety of different parameters which are formatted differently.

<details>
  <summary>Expand for the code used to produce the example data</summary>
```{r}
dat <- tibble::tribble(
  ~label     , ~param, ~column    , ~ value,   ~ord, 
  "n"        ,"n"     , "trt1"     ,12,        1, 
  "mean (sd)","mean"  , "trt1"     ,12.332837, 2,
  "mean (sd)","sd"    , "trt1"     ,4.3454547, 2,
  "median"   ,"median", "trt1"     ,14,        3, 
  "[q1, q3]" ,"q1"    , "trt1"     ,10,        4,
  "[q1, q3]" ,"q3"    , "trt1"     ,20,        4,
  "n"        ,"n"     , "trt2"     ,24,        1,
  "mean (sd)","mean"  , "trt2"     ,15.438737, 2,
  "mean (sd)","sd"    , "trt2"     ,6.723827,  2, 
  "median"   ,"median", "trt2"     ,16,        3, 
  "[q1, q3]" ,"q1"    , "trt2"     ,11,        4, 
  "[q1, q3]" ,"q3"    , "trt2"     ,22,        4,
  "n"        ,"pval"  , "p-value"  ,NA,        1, 
  "mean (sd)","pval"  , "p-value"  ,0.00002,   2, 
  "median"   ,"pval"  , "p-value"  ,0.051211,  3, 
  "[q1, q3]" ,"pval"  , "p-value"  ,NA,        4)
```
</details>



```{r}
head(dat)

tfrmt(
  label = label,
  column = column,
  param = param,
  value = value,
  sorting_cols = c(ord),
  col_plan = col_plan(-ord), 
  body_plan = body_plan(
    frmt_structure(group_val = ".default", label_val = ".default", 
                   frmt("xx", missing = " ")),
    frmt_structure(group_val = ".default", label_val = ".default", 
                   frmt_combine("{mean} ({sd})", 
                                mean = frmt("xx.x"), 
                                sd = frmt("xx.xx"),
                                missing=" ")),
    frmt_structure(group_val = ".default", label_val = ".default", 
                   frmt_combine("[{q1}, {q3}]", 
                                q1 = frmt("xx.x"), 
                                q3 = frmt("xx.x"),
                                missing=" ")),
    frmt_structure(group_val = ".default", label_val = ".default", 
                   pval = frmt_when(">0.99" ~ ">0.99",
                                    "<0.001" ~ "<0.001",
                                    "<0.05" ~ frmt("x.xxx*"),
                                    TRUE ~ frmt("x.xxx", missing = "")))
  )
) %>% 
  print_to_gt(dat)
```

Column alignment can improve this table by making it easier to read. Let's start by left-aligning our p-value column. Notice that by providing a `col_style_plan`, any variables that are not covered in the plan will be left-aligned by default.

```{r}
tfrmt(
  label = label,
  column = column,
  param = param,
  value = value,
  sorting_cols = c(ord),
  col_plan = col_plan(-ord), 
  body_plan = body_plan(
    frmt_structure(group_val = ".default", label_val = ".default", 
                   frmt("xx", missing = " ")),
    frmt_structure(group_val = ".default", label_val = ".default", 
                   frmt_combine("{mean} ({sd})", 
                                mean = frmt("xx.x"), 
                                sd = frmt("xx.xx"),
                                missing=" ")),
    frmt_structure(group_val = ".default", label_val = ".default", 
                   frmt_combine("[{q1}, {q3}]", 
                                q1 = frmt("xx.x"), 
                                q3 = frmt("xx.x"),
                                missing=" ")),
    frmt_structure(group_val = ".default", label_val = ".default", 
                   pval = frmt_when(">0.99" ~ ">0.99",
                                    "<0.001" ~ "<0.001",
                                    "<0.05" ~ frmt("x.xxx*"),
                                    TRUE ~ frmt("x.xxx", missing = "")))
  ), 
  col_style_plan =  col_style_plan(
    col_style_structure(col = `p-value`, align = "left", type = "char")
  )
  ) %>% 
  print_to_gt(dat)
```

This alignment isn't quite right as we have the `<` in one value but not the other. Applying a decimal alignment would be a better fit:

```{r}
tfrmt(
  label = label,
  column = column,
  param = param,
  value = value,
  sorting_cols = c(ord),
  col_plan = col_plan(-ord), 
  body_plan = body_plan(
    frmt_structure(group_val = ".default", label_val = ".default", 
                   frmt("xx", missing = " ")),
    frmt_structure(group_val = ".default", label_val = ".default", 
                   frmt_combine("{mean} ({sd})", 
                                mean = frmt("xx.x"), 
                                sd = frmt("xx.xx"),
                                missing=" ")),
    frmt_structure(group_val = ".default", label_val = ".default", 
                   frmt_combine("[{q1}, {q3}]", 
                                q1 = frmt("xx.x"), 
                                q3 = frmt("xx.x"),
                                missing=" ")),
    frmt_structure(group_val = ".default", label_val = ".default", 
                   pval = frmt_when(">0.99" ~ ">0.99",
                                    "<0.001" ~ "<0.001",
                                    "<0.05" ~ frmt("x.xxx*"),
                                    TRUE ~ frmt("x.xxx", missing = "")))
  ), 
  col_style_plan =  col_style_plan(
    col_style_structure(col = `p-value`, align = c("."), type = "char"))
  ) %>% 
  print_to_gt(dat)
```

For our other two columns, we have a mix of values represented. In this case, we want to align on the first set of digits. In other words, we will align on the first instance of a decimal, comma, or space:

```{r}
tfrmt(
  label = label,
  column = column,
  param = param,
  value = value,
  sorting_cols = c(ord),
  col_plan = col_plan(-ord), 
  body_plan = body_plan(
    frmt_structure(group_val = ".default", label_val = ".default", 
                   frmt("xx", missing = " ")),
    frmt_structure(group_val = ".default", label_val = ".default", 
                   frmt_combine("{mean} ({sd})", 
                                mean = frmt("xx.x"), 
                                sd = frmt("xx.xx"),
                                missing=" ")),
    frmt_structure(group_val = ".default", label_val = ".default", 
                   frmt_combine("[{q1}, {q3}]", 
                                q1 = frmt("xx.x"), 
                                q3 = frmt("xx.x"),
                                missing=" ")),
    frmt_structure(group_val = ".default", label_val = ".default", 
                   pval = frmt_when(">0.99" ~ ">0.99",
                                    "<0.001" ~ "<0.001",
                                    "<0.05" ~ frmt("x.xxx*"),
                                    TRUE ~ frmt("x.xxx", missing = "")))
  ), 
  col_style_plan =  col_style_plan(
    col_style_structure(col = `p-value`, align = c("."), type = "char"),
    col_style_structure(col = starts_with("trt"), align = c(".", ",", " "), type = "char")
    )
  ) %>% 
  print_to_gt(dat)
```

Finally, for the purpose of demonstrating our options, we can align each column differently (left, right, character for the three columns respectively):

```{r}
tfrmt(
  label = label,
  column = column,
  param = param,
  value = value,
  sorting_cols = c(ord),
  col_plan = col_plan(-ord), 
  body_plan = body_plan(
    frmt_structure(group_val = ".default", label_val = ".default", 
                   frmt("xx", missing = " ")),
    frmt_structure(group_val = ".default", label_val = ".default", 
                   frmt_combine("{mean} ({sd})", 
                                mean = frmt("xx.x"), 
                                sd = frmt("xx.xx"),
                                missing=" ")),
    frmt_structure(group_val = ".default", label_val = ".default", 
                   frmt_combine("[{q1}, {q3}]", 
                                q1 = frmt("xx.x"), 
                                q3 = frmt("xx.x"),
                                missing=" ")),
    frmt_structure(group_val = ".default", label_val = ".default", 
                   pval = frmt_when(">0.99" ~ ">0.99",
                                    "<0.001" ~ "<0.001",
                                    "<0.05" ~ frmt("x.xxx*"),
                                    TRUE ~ frmt("x.xxx", missing = "")))
  ), 
  col_style_plan =  col_style_plan(
    col_style_structure(col = `p-value`, align = c("."), type = "char"),
    col_style_structure(col = trt1, align = "left", type = "char"),
    col_style_structure(col = trt2, align = "right", type = "char"))
  ) %>% 
  print_to_gt(dat)
```

### Alignment Examples - Positional

For the ability to align across multiple characters and/or positions, positional alignment gives us the most flexibility. As such, it also requires a more granular representation of how we wish to align our values. For instance, instead of specifying an alignment on the first decimal: `align = "."`, we provide a character vector representation of our formatted cells with vertical bars placed in any position we wish to align: `align = c("xx|.x", "x|.x")`. 

In our example from above, we can use this feature to align on multiple decimal places:
```{r}
tf <- tfrmt(
  label = label,
  column = column,
  param = param,
  value = value,
  sorting_cols = c(ord),
  col_plan = col_plan(-ord), 
  body_plan = body_plan(
    frmt_structure(group_val = ".default", label_val = ".default", 
                   frmt("xx", missing = " ")),
    frmt_structure(group_val = ".default", label_val = ".default", 
                   frmt_combine("{mean} ({sd})", 
                                mean = frmt("xx.x"), 
                                sd = frmt("xx.xx"),
                                missing=" ")),
    frmt_structure(group_val = ".default", label_val = ".default", 
                   frmt_combine("[{q1}, {q3}]", 
                                q1 = frmt("xx.x"), 
                                q3 = frmt("xx.x"),
                                missing=" ")),
    frmt_structure(group_val = ".default", label_val = ".default", 
                   pval = frmt_when(">0.99" ~ ">0.99",
                                    "<0.001" ~ "<0.001",
                                    "<0.05" ~ frmt("x.xxx*"),
                                    TRUE ~ frmt("x.xxx", missing = "")))
  ), 
  col_style_plan =  col_style_plan(
    col_style_structure(col = `p-value`, align = c("."), type = "char"),
    col_style_structure(col = starts_with("trt"), 
                        align = c("xx|",
                                  "xx|.x ( x|.xx)",
                                  "[xx|.x, xx|.x]"), 
                        type = "pos")
  )
) 

tf %>%  print_to_gt(dat)
```

Behind the scenes, {tfrmt} makes a few assumptions about how these vertical bars translate to alignment:

- Any characters to the left of the first "|" will be left-padded as needed. 
- Any characters to the right of the final "|" will be right-padded as needed
- Any characters between two "|"s will receive padding that is anchored to the first occurance of whitespace in that substring (e.g. "xx|.x (xx.|x)" will be padded between the ".x" and "(xx.").)
    - If no whitespace is found, no padding will be applied to that portion of the string

It is important to note that if we are not creating a mock table, the `align` input should cover scenarios observed in the provided data. There is a helper function called `display_val_frmts` that helps to create the `align` input based on the `tfrmt` object (specifically its `body_plan`) and the provided data. The output of this helper function is a a bit of code that can be copied and modified by adding the vertical bars as desired:
```{r}
display_val_frmts(tfrmt = tf, .data = dat, col = vars(starts_with("trt")))
```

### Alignment of group and label columns

At this time, group and label columns will be left aligned by default. The user can adjust the alignment of these columns through gt directly.

### Alignment via `body_plan`

It should be noted that in simple cases, the user can achieve desired alignment via the `frmt`s in the `body_plan`. These `frmt` specifications  allow us to pad values to a desired length. For example, `frmt(xxx)` will ensure values with fewer than 3 digits will be padded accordingly: 23 will become " 23" and 9 will become "  9". In a straightforward table of values with similar scales, this may be sufficient. However in many cases the user may desire more control over alignment and the `col_style_plan` is the preferred method. 

## Width

Width is set as specific value in number of characters wide the column is allowed to be at most. This value is specified by passing a character or numeric value to the "width" argument of `col_style_structure`, taking the format of "##" or ##, where "##" is the number of characters.

Width can be applied to any of the columns in the table.

### Example

Continuing the example from alignment, we set the widths of the columns in the table. Because "label" is output in the table, we can apply the widths to that column as well.

```{r}
tfrmt(
  label = label,
  column = column,
  param = param,
  value = value,
  sorting_cols = c(ord),
  col_plan = col_plan(-ord), 
  body_plan = body_plan(
    frmt_structure(group_val = ".default", label_val = ".default", 
                   frmt("xx", missing = " ")),
    frmt_structure(group_val = ".default", label_val = ".default", 
                   frmt_combine("{mean} ({sd})", 
                                mean = frmt("xx.x"), 
                                sd = frmt("xx.xx"),
                                missing=" ")),
    frmt_structure(group_val = ".default", label_val = ".default", 
                   frmt_combine("[{q1}, {q3}]", 
                                q1 = frmt("xx.x"), 
                                q3 = frmt("xx.x"),
                                missing=" ")),
    frmt_structure(group_val = ".default", label_val = ".default", 
                   pval = frmt_when(">0.99" ~ ">0.99",
                                    "<0.001" ~ "<0.001",
                                    "<0.05" ~ frmt("x.xxx*"),
                                    TRUE ~ frmt("x.xxx", missing = "")))
  ), 
  col_style_plan =  col_style_plan(
    col_style_structure(col = label, width = "10"),
    col_style_structure(col = `p-value`, align = c("."), width = "10"),
    col_style_structure(col = trt1, align = "left", width = "5"),
    col_style_structure(col = trt2, align = "right", width = "5"))
  ) %>% 
  print_to_gt(dat)

```
