---
title: "Extract Data from a tfrmt"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Extract Data from a tfrmt}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, message = FALSE, warning = FALSE}
library(tfrmt)
library(dplyr)
library(gt)
```

The `{tfrmt}` package is designed to create highly formatted Tables by layering formatting logic onto data. While the primary output of `print_to_gt()` is a `gt` object ready for reporting, there are times when you may need to access the "final" processed data frame—for example, to perform secondary validations, to export the data, or use in another reporting tool.

The `extract_data()` function provides a clean way to pull this underlying data back out of a `gt_tbl` or `gt_group` object. `extract_data()` handles the cleanup of the data, removing internal columns and ensuring that column names reflect the labels seen in the final table.

## Basic Usage

In this example, we take a demographic dataset, apply a tfrmt, and then extract the resulting data frame.

```{r}
# Subset data for a simple example
data_demog_test <- data_demog |>
  filter(
    rowlbl1 %in% c("Age (y)", "Sex"),
    column != "p-value"
  )

# Define and print the tfrmt
tfrmt_obj <- tfrmt(
  group = c(rowlbl1, grp),
  label = rowlbl2,
  column = column,
  param = param,
  value = value,
  sorting_cols = c(ord1, ord2),
  # specify value formatting
  body_plan = body_plan(
    frmt_structure(group_val = ".default", label_val = ".default", frmt_combine("{n} {pct}",
      n = frmt("xxx"),
      pct = frmt_when(
        "==100" ~ "",
        "==0" ~ "",
        TRUE ~ frmt("(xx.x %)")
      )
    )),
    frmt_structure(group_val = ".default", label_val = "n", frmt("xxx")),
    frmt_structure(group_val = ".default", label_val = c("Mean", "Median", "Min", "Max"), frmt("xxx.x")),
    frmt_structure(group_val = ".default", label_val = "SD", frmt("xxx.xx"))
  ),
  # remove extra cols
  col_plan = col_plan(
    -grp,
    -ord1,
    -ord2
  )
)

gt_table <- print_to_gt(tfrmt_obj, data_demog_test)

gt_table

# Extract the data
extracted_df <- extract_data(gt_table)

extracted_df
```


## Extracting from Paged Tables (`gt_group`)

When using a `page_plan`, `print_to_gt()` returns a `gt_group` object containing multiple tables. `extract_data()` automatically detects this and returns a list of data frames, one for each page.

In this example we add pagination to our table using `page_plan()` and extract the data from the `gt_group` object using `extract_data()`.

```{r}
tfrmt_paged <- tfrmt_obj |>
  tfrmt(
    page_plan = page_plan(
      page_structure(
        group_val = list(
          rowlbl1 = ".default"
        )
      ),
      note_loc = "source_note"
    )
  )

gt_paged <- print_to_gt(tfrmt_paged, data_demog_test)

gt_paged

# This returns a list of data frames
data_list <- extract_data(gt_paged)

data_list
```


## Integration with Big N

`extract_data()` is also compatible with `big_n_structure`. If your table includes Big N values in the column headers, these formatted strings will be preserved as the column names in the extracted data frame.

Example showing big Ns preserved in the column names after using `extract_data()`.

```{r}
data <- tibble::tibble(
  Group = c("N", "N", "N", rep(c("Age (y)", "Sex", "Age (y)", "Sex"), c(3, 3, 6, 12))),
  Label = c("N", "N", "N", rep(c("n", "Mean (SD)", "Male", "Female"), c(6, 6, 6, 6))),
  Column = c("Placebo", "Treatment", "Total", rep(c("Placebo", "Treatment", "Total"), times = 8)),
  Param = c("bigN", "bigN", "bigN", rep(c("n", "mean", "sd", "n", "pct", "n", "pct"), c(6, 3, 3, 3, 3, 3, 3))),
  Value = c(
    30, 40, 60, 15, 13, 28, 14, 13, 27, 73.56, 74.231, 71.84, 9.347, 7.234, 8.293,
    8, 7, 15, 8 / 14, 7 / 13, 15 / 27, 6, 6, 12, 6 / 14, 6 / 13, 12 / 27
  )
) |>
  dplyr::mutate(
    Value = dplyr::case_when(
      Param == "pct" ~ Value * 100,
      TRUE ~ Value
    ),
    ord1 = dplyr::if_else(Param == "bigN", 0, 1),
    ord2 = dplyr::if_else(Param == "bigN", 0, 1)
  )


bign <- tfrmt(
  group = Group,
  label = Label,
  column = Column,
  value = Value,
  param = Param,
  sorting_cols = c(ord1, ord2),
  body_plan = body_plan(
    frmt_structure(
      group_val = ".default",
      label_val = ".default",
      frmt_combine("{n} {pct}",
        n = frmt("X"),
        pct = frmt("(xx.x%)", missing = " ")
      )
    ),
    frmt_structure(
      group_val = "Age (y)", label_val = "Mean (SD)",
      frmt_combine("{mean} ({sd})",
        mean = frmt("XX.X"),
        sd = frmt("x.xx")
      )
    ),
    frmt_structure(group_val = ".default", label_val = "n", frmt("xx"))
  ),
  col_plan = col_plan(everything(), -starts_with("ord"), "Total"),
  row_grp_plan = row_grp_plan(
    row_grp_structure(group_val = ".default", element_block(post_space = " "))
  ),
  big_n = big_n_structure(param_val = "bigN", n_frmt = frmt("\nN = xx"))
) |>
  print_to_gt(data)

bign

extracted <- extract_data(bign)

extracted
```

## Handling Spanning Headers

If your table has multiple levels of column headers, `gt` stores these internally using a delimiter.

By default, `extract_data()` will collapse these levels using an underscore (`_`), but you can specify a custom delimiter using the `col_delim` argument.

Example showing how spanning headers are combined using `col_delim` argument in `extract_data()`.

```{r}
data <- tibble::tribble(
  ~group, ~label, ~span2, ~span1, ~my_col, ~parm, ~val,
  "g1", "rowlabel1", "column cols", "cols 1,2", "col1", "value", 1,
  "g1", "rowlabel1", "column cols", "cols 1,2", "col2", "value", 1,
  "g1", "rowlabel1", NA, NA, "mycol3", "value", 1,
  "g1", "rowlabel1", "column cols", "col 4", "col4", "value", 1,
  "g1", "rowlabel1", NA, NA, "mycol5", "value", 1,
  "g1", "rowlabel2", "column cols", "cols 1,2", "col1", "value", 2,
  "g1", "rowlabel2", "column cols", "cols 1,2", "col2", "value", 2,
  "g1", "rowlabel2", NA, NA, "mycol3", "value", 2,
  "g1", "rowlabel2", "column cols", "col 4", "col4", "value", 2,
  "g1", "rowlabel2", NA, NA, "mycol5", "value", 2,
  "g2", "rowlabel3", "column cols", "cols 1,2", "col1", "value", 3,
  "g2", "rowlabel3", "column cols", "cols 1,2", "col2", "value", 3,
  "g2", "rowlabel3", NA, NA, "mycol3", "value", 3,
  "g2", "rowlabel3", "column cols", "col 4", "col4", "value", 3,
  "g2", "rowlabel3", NA, NA, "mycol5", "value", 3,
)

# 2 layers of spanning headers
spanning_tfrmt <- tfrmt(
  group = group,
  label = label,
  param = parm,
  value = val,
  column = c(span2, span1, my_col),
  body_plan = body_plan(
    frmt_structure(group_val = ".default", label_val = ".default", frmt("x"))
  ),
  col_plan = col_plan(
    group,
    label,
    starts_with("col")
  )
) |> print_to_gt(data)

spanning_tfrmt

res_layer <- extract_data(spanning_tfrmt, col_delim = "/")
res_layer
```
