---
title: "FAQ"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{FAQ}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, message=FALSE, warning=FALSE}
library(tfrmt)
library(dplyr)
library(gt)
```

### Can you format a column header?

To format your column header, just use markdown syntax within your character strings. Examples of this could be a newline, or bolding certain words. To add a newline to your column header you need to include the markdown syntax `<br>` or `\n` in your character string. For bolding, surround your text with `**` on either side. Some example code is provided below:

```{r}
es_data <-tibble(rowlbl1 =c(rep("Completion Status",12),rep("Primary reason for withdrawal",28)),
                 rowlbl2 =c(rep("Completed",4),rep("Prematurely Withdrawn",4),rep("Unknown",4),rep("Adverse Event",4),rep("Lost to follow-up",4),rep("Protocol violation",4),rep("Subject decided to withdraw",4),rep("Protocol Violation",4),rep("Pre-Operative Dose[1]",4),rep("Other",4)),
                 param=c(rep(c("n","n","pct","pct"),10)),
                 column=c(rep(c("Placebo<br>(N=48)","Treatment\n**(N=38)**"),20)), #newline and bold syntax
                 value=c(24,19,2400/48,1900/38,5,1,500/48,100/38,19,18,1900/48,1800/38,1,1,100/48,100/38,0,0,0,0,0,0,0,0,1,1,100/48,100/38,1,4,100/48,400/38,1,0,100/48,0,2,3,200/48,300/38)

)

tfrmt(
  # specify columns in the data
  group = c(rowlbl1),
  label = rowlbl2,
  column = column,
  param = param,
  value = value,
  # set formatting for values
  body_plan = body_plan(
    frmt_structure(
      group_val = ".default",
      label_val = ".default",
      frmt_combine(
        "{n} {pct}",
        n = frmt("xxx"),
        pct = frmt_when("==100" ~ "",
                        "==0" ~ "",
                        TRUE ~ frmt("(xx.x %)"))
      )
    )
  ),
  
  # Specify row group plan
  # Indent the rowlbl2
  row_grp_plan = row_grp_plan(
    row_grp_structure(group_val = ".default", element_block(post_space = " ")),
    label_loc = element_row_grp_loc(location = "indented")
  )
) %>%
  print_to_gt(es_data) %>%
  tab_options(container.width = 1000)
```

### How do I output my table?

To output your table, use `tfrmt`'s `print_to_gt` and `print_mock_gt` functions, which create `gt` table objects. We recommend leveraging the export capabilities provided by the `gt` package to save and share your table. For more information on exporting, see [the gt package documentation](https://gt.rstudio.com/reference/gtsave.html).
