---
title: "Using Footnotes"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Using Footnotes}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, message = FALSE, warning = FALSE}
library(tfrmt)
library(dplyr)
library(gt)
library(tidyr)
```

In `tfrmt` footnotes are a table component which can be added and used to the 
specification in a manner similar to `body_plan` and `col_plan`. It has its own 
dedicated function called `footnote_plan` which consists of a list of 
`footnote_structure`s.

Some options for the marks used to indicate footnotes are the following:
 - letters (letters: a, b, c, etc.)
 - numbers (numbers: 1, 2, 3, etc.)
 - standard (columns: asterisk, groups: obelisk/dagger)

# Example (with groups)

```{r}
# Create data
df <- crossing(group = c("group 1", "group 2"),
               label = c("label 1", "label 2"),
               column = c("PL", "T1", "T2", "T1&T2"),
               param = c("count", "percent"))

# Write specification
tfrmt(
  group = group,
  label = label,
  column = column,
  param = param,
  row_grp_plan = row_grp_plan(
    row_grp_structure(group_val = ".default",
                      element_block(post_space = "   ")) ),
  
  body_plan = body_plan(
    frmt_structure(group_val = ".default", label_val = ".default",
                   frmt_combine("{count} ({percent})",
                                count = frmt("xx"),
                                percent = frmt("xx.x")))
  ),
  
  # Add footnotes here:
  footnote_plan = footnote_plan(
    footnote_structure(footnote_text = "Source Note", group_val = "group 1"),
    footnote_structure(footnote_text = "Placebo", column_val = "PL"),
    marks = "numbers"),
  
  ) %>%
  print_mock_gt(df)
```

