---
title: "JSON metadata"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{JSON metadata}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, message = FALSE, warning = FALSE}
library(tfrmt)
```

Aside from R code, `tfrmt` objects can be represented as machine-readable metadata in `JSON` format. This enables users to save formatting metadata to a database, or alternatively import templates stored as `JSON` files. There are several utilities available for seamless translation between `tfrmt` and `JSON`. 

## tfrmt to JSON

A `tfrmt` object can be converted to `JSON` using the `tfrmt_to_json` function. Suppose we have a basic `tfrmt`:

```{r}
template_tfrmt <- tfrmt(
  group = grp,
  label = label,
  column = column,
  param = param,
  value = value,
  body_plan = body_plan(
    frmt_structure(
      group_val = ".default",
      label_val = ".default",
      frmt("xx.x")
    )
  )
)
template_tfrmt %>% print_mock_gt()
```

We can pass this through `tfrmt_to_json` to convert to `JSON` metadata. 
```{r}
template_tfrmt %>%
  tfrmt_to_json()
```

This `JSON` can optionally be saved to a file by providing a file path as the second argument to the function:

```{r eval = FALSE}
tfrmt(
  group = grp,
  label = label,
  column = column,
  param = param,
  value = value,
  body_plan = body_plan(
    frmt_structure(
      group_val = ".default",
      label_val = ".default",
      frmt("xx.x")
    )
  )
) %>%
  tfrmt_to_json(path = "template.JSON")
```


## JSON to tfrmt

If a `JSON` file already exists, users can import this into R.

```{r echo=FALSE}
template_json <- tfrmt(
  group = grp,
  label = label,
  column = column,
  param = param,
  value = value,
  title = "mytitle",
  body_plan = body_plan(
    frmt_structure(
      group_val = ".default",
      label_val = ".default",
      frmt("xx.x")
    )
  )
)
```

```{r eval = FALSE}
template_json <- json_to_tfrmt(path = "template.JSON")
```
 

Once available in the R session, users can optionally build on this template as needed.
```{r}
template_json %>%
  layer_tfrmt(
    tfrmt(title = "Custom title")
  ) %>%
  print_mock_gt()
```


