---
title: "Mock Tables"
author: "Karima Ahmad"
date: "10/06/2022"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Mock Tables}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, message=FALSE}
library(tlang)
library(dplyr)
library(tidyr)
```

## AE Mocks

Here is some mock data created using the `crossing()` function from the `tidyr` package.

```{r mock data}
df <- crossing(label = c("label 1", "label 2", "label 3"),
               column = c("trt1", "trt2", "trt1&trt2", "pl"),
               param = c("count", "percent")) %>%
  # Assign numerical order (optional)
  mutate(ord1 = rep(seq(1:length(unique(.$label))), each = nrow(.)/length(unique(.$label)) ))

df
```

These mock tables don't contain numerical values as none have been supplied to 
the tfrmt specification. If your dataset contains values that you would like to 
add then pass an argument to the `value` parameter when applying the `tfrmt()` 
function.

Note: Layering has been used to avoid repeating code; for more information 
please see the "Reuse and layer tfrmts" article.


```{r}
mock_tbl <- function(tfrmt_obj){
  tfrmt(
    # Specify columns in the data
    label = label,
    column = column, 
    param = param,
    sorting_cols = c(ord1),
    
    # Specify body plan
    body_plan = body_plan(
      frmt_structure(group_val = ".default", label_val = ".default",
                     frmt_combine(
                       "{count} {percent}",
                       count = frmt("XXX"),
                       percent = frmt_when("==100"~ frmt(""),
                                           "==0"~ "",
                                           "TRUE" ~ frmt("(XX.X%)"))))
      ))
}

# Print table
mock_tbl() %>%
  print_mock_gt()
```

You can adjust column names and structure using the `col_plan` parameter. It 
accepts selection helpers from the `tidyselect` package. For example, you can 
select `everything()` and remove columns that `starts_with("ord")`.

To order and rename columns:

```{r}
# this example is currently not working...

mock_tbl_adjust_cols <- function(){
  mock_tbl() %>%
  tfrmt(
    # Specify column structure
    col_plan = col_plan(
      -starts_with("ord"),
      span_structure(
        "Treatment",
        T1 = trt1,
        T2 = trt2,
        `T1&T2`= `trt1&trt2`
        ),
      span_structure(
        "Placebo",
        PL = pl
        )
      
      )
  ) 
}
  
mock_tbl_adjust_cols() %>%
  print_mock_gt(df)
```

To add a title, subtitle or footer:

```{r}
mock_tbl_add_titles <- function(){
  mock_tbl_adjust_cols() %>%
    tfrmt(
      # Specify title, subtitle, footer
      title = "Table Name",
      subtitle = "Study ID: GSK12345",
      footer = "A footnote about stuff<br/>Here is a new line with more stuff")
}
  
mock_tbl_add_titles() %>%
  print_mock_gt(df)

```

Mock table with combined columns:

```{r}
fmt_spec <- tfrmt(
  group = AEBODSYS,
  label = AETERM,
  param = param,
  column = c(col2, col1),
  value = value,
  row_grp_plan = row_grp_plan(),
  body_plan = body_plan(
    frmt_structure(group_val = ".default", label_val = ".default",
            frmt_combine("{n} {pct}",
                        n = frmt("XXX"),
                        pct = frmt_when(
                          "==100" ~ "",
                          "==0" ~ "",
                          TRUE ~ frmt("(xx.x %)")))),
    frmt_structure(group_val = ".default", label_val = ".default", 
                   AEs = frmt("[XXX]")),
    frmt_structure(group_val = ".default", label_val = ".default", 
                   pval = frmt_when(">0.99" ~ ">0.99",
                                    "<0.001" ~ "<0.001",
                                    "<0.05" ~ frmt("x.xxx*"),
                                    TRUE ~ frmt("x.xxx", missing = "--")))
  ), col_plan = col_plan(-starts_with("ord")))

ae_data2 %>%  
  select(-value) %>% 
  arrange(ord1, ord2 ) %>% 
  print_mock_gt(fmt_spec, . )  %>% 
  tab_options(
    container.width = 1000
  )
```


### AE Mock Table Examples

Reusable tfrmt base layer:

```{r}
ae_mock_table <- function(tfrmt_obj){
  tfrmt(
    title = "Table Name",
    subtitle = "Study ID: GSK12345",
    footer = "A footnote about stuff<br/>Here is a new line with more stuff",  body_plan = body_plan(
    frmt_structure(group_val = ".default", label_val = ".default",
                   frmt_combine(
                     "{count} {percent}",
                     count = frmt("XXX"),
                     percent = frmt_when("==100"~ frmt(""),
                                         "==0"~ "",
                                         "TRUE" ~ frmt("(XX.X%)")))),
    frmt_structure(group_val = ".default", label_val = ".default",
                   frmt_combine(
                     "{num} ({lower}, {upper})",
                     num = frmt("XX.X"),
                     lower = frmt_when("==100"~ frmt(""),
                                       "==0"~ "",
                                       "TRUE" ~ frmt("XX.X%")),
                     upper = frmt_when("==100"~ frmt(""),
                                       "==0"~ "",
                                       "TRUE" ~ frmt("XX.X%"))))
    ),
  col_plan = col_plan(
    T1, T2, `T1&T2`, PL,
    everything(),
    -starts_with("ord")
    )
  )
}
```

Example 1:

```{r}
df <- bind_rows(
  crossing(label = c("label 1", "label 2", "label 3"),
           column = c("T1", "T2", "T1&T2", "PL"),
           param = c("count", "percent")) %>%
    mutate(ord1 = rep(seq(1:length(unique(.$label))), each = nrow(.)/length(unique(.$label)) )),
  crossing(label = c("label 1", "label 2", "label 3"),
           column = c("risk T1-PL", "risk T2-PL", "risk T1&T2-PL"),
           param = c("num", "lower", "upper")) %>%
    mutate(ord1 = rep(seq(1:length(unique(.$label))), each = nrow(.)/length(unique(.$label)) ))
  ) %>% arrange_all()

ae_mock_table() %>%
  tfrmt(
    label = "label",
    param = "param",
    column = "column",
    sorting_cols = vars(ord1)
    ) %>%
  print_mock_gt(df)
```

Example 2:

```{r}
df <- bind_rows(
  crossing(group = c("grp1", "grp2"),
           label = c("label 1", "label 2", "label 3"),
           column = c("T1", "T2", "T1&T2", "PL"),
           param = c("count", "percent")),
  crossing(group = c("grp1", "grp2"),
           label = c("label 1", "label 2", "label 3"),
           column = c("risk T1-PL", "risk T2-PL", "risk T1&T2-PL"),
           param = c("num", "lower", "upper"))
  ) %>% arrange_all()

ae_mock_table() %>%
  tfrmt(
    group = group,
    label = "label",
    param = "param",
    column = "column",
    row_grp_plan = row_grp_plan( row_grp_structure(group_val = ".default", element_block(post_space = "   ")) ),
    
    col_plan = col_plan(
      T1, T2, `T1&T2`, PL, `risk T1-PL`, `risk T2-PL`, `risk T1&T2-PL`,
      everything(),
      -starts_with("ord"))
    ) %>%
  print_mock_gt(df)

```

Example 3:

```{r}
df <- bind_rows(
  crossing(grp1 = c("group 1.1", "group 1.2"),
           grp2 = c("group 2.1", "group 2.2"),
           label = c("label 1", "label 2"),
           column = c("T1", "T2", "T1&T2", "PL"),
           param = c("count", "percent")),
  crossing(grp1 = c("group 1.1", "group 1.2"),
           grp2 = c("group 2.1", "group 2.2"),
           label = c("label 1", "label 2"),
           column = c("risk T1-PL", "risk T2-PL", "risk T1&T2-PL"),
           param = c("num", "lower", "upper"))
  ) %>% arrange_all()

ae_mock_table() %>%
  tfrmt(
    group = c(grp1, grp2),
    label = "label",
    param = "param",
    column = "column",
    row_grp_plan = row_grp_plan( row_grp_structure(group_val = ".default", element_block(post_space = "   ")) ),
    
    col_plan = col_plan(
      T1, T2, `T1&T2`, PL, `risk T1-PL`, `risk T2-PL`, `risk T1&T2-PL`,
      everything(),
      -starts_with("ord"))
    ) %>%
  print_mock_gt(df)

```

Example 4:

```{r}
ae_mock_table() %>%
  tfrmt(
    group = c(grp1, grp2),
    label = "label",
    param = "param",
    column = "column",
    row_grp_plan = row_grp_plan(label_loc = element_row_grp_loc(location = "column")),    
    
    col_plan = col_plan(
      T1, T2, `T1&T2`, PL, `risk T1-PL`, `risk T2-PL`, `risk T1&T2-PL`,
      everything(),
      -starts_with("ord"))
    ) %>%
  print_mock_gt(df)

```


## Demography Mocks

```{r}
tfrmt(
  # specify columns in the data
  group = c(rowlbl1,grp),
  label = rowlbl2,
  column = column, 
  param = param,
  values = value,
  sorting_cols = c(ord1, ord2),
  
  # Specify body plan
  body_plan = body_plan(
    frmt_structure(group_val = ".default", label_val = ".default", frmt_combine("{n} {pct}", 
                                                                                n = frmt("xxx"),
                                                                                pct = frmt_when("==100" ~ "",
                                                                                                "==0" ~ "",
                                                                                                TRUE ~ frmt("(xx.x %)")))),
    frmt_structure(group_val = ".default", label_val = "n", frmt("xxx")),
    frmt_structure(group_val = ".default", label_val = c("Mean", "Median", "Min","Max"), frmt("xxx.x")),
    frmt_structure(group_val = ".default", label_val = "SD", frmt("xxx.xx")),
    frmt_structure(group_val = ".default", label_val = ".default", p = frmt("")),
    frmt_structure(group_val = ".default", label_val = c("n","<65 yrs","<12 months","<25"), p = frmt_when(">0.99" ~ ">0.99",
                                                                                 "<0.001" ~ "<0.001",
                                                                                 TRUE ~ frmt("x.xxx", missing = "")))
  ),
  
  # Specify row group plan
  row_grp_plan = row_grp_plan(
    row_grp_structure(group_val = ".default", element_block(post_space = " ")),
    label_loc = element_row_grp_loc(location = "column")
  ),
  
  # Specify column alignment plan
  col_align = col_align_plan(
    element_align(align = c(".",","," "), col = vars(everything()))
  ),
  
  # remove extra cols
  col_plan = col_plan(-grp, 
                      -starts_with("ord") )
) %>% 
  print_mock_gt(demog_data %>% select(-value)) %>% 
  tab_options(
    container.width = 900
  )
```


## Efficacy Mocks

```{r}
tfrmt(
  group = group,
  label = label,
  column = column,
  param = param,
  values = value,
  sorting_cols = c(ord1, ord2),
  row_grp_plan = row_grp_plan(
    row_grp_structure(group_val = list(group="Change from Baseline"), element_block(post_space = " ")),
    row_grp_structure(group_val = list(group="p-value (Dose Response)"), element_block(post_space = " ")),
    row_grp_structure(group_val = list(group="p-value (Xan - Placebo)"), element_block(post_space = " "))
  ),
  body_plan = body_plan(
    frmt_structure(group_val = ".default", label_val = "n", frmt("xx")), 
    frmt_structure(group_val = ".default", label_val = "Median (Range)", frmt_combine("{median} ({min};{max})",
                                                                                      median = frmt("xx.x"),
                                                                                      min = frmt("xx"),
                                                                                      max = frmt("xx"), missing = " ")),
    frmt_structure(group_val = ".default", label_val = "Mean (SD)", frmt_combine("{mean} ({sd})",
                                                                                      mean = frmt("xx.x"),
                                                                                      sd = frmt("xx.xx"), missing = " ")),
    frmt_structure(group_val = ".default", label_val = "Diff of LS Means (SE)", frmt_combine("{diff} ({diff_se})",
                                                                                 diff = frmt("xx.x"),
                                                                                 diff_se = frmt("xx.xx"), missing = " ")),
    frmt_structure(group_val = ".default", label_val = "95% CI", frmt_combine("({diff_lcl};{diff_ucl})",
                                                                              diff_lcl = frmt("xx.x"),
                                                                              diff_ucl = frmt("xx.x"), missing = " ")),
    frmt_structure(group_val = ".default", label_val = ".default", p.value = frmt_when("<0.001" ~ "<0.001",
                                                                                       ">0.99" ~ ">0.99",
                                                                                       TRUE ~ frmt("x.xxx", missing = " ")))
  ),
  col_plan = col_plan(
    group, label,
    contains("Placebo"),
    contains("Low"),
    contains("High")
  )
) %>%
  print_mock_gt(efficacy_data %>% select(-value))%>% 
  tab_options(
    container.width = 800
  )
```



