---
title: "Mock Table Examples"
author: "Karima Ahmad"
date: "10/06/2022"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Mock Tables}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, message=FALSE}
library(tlang)
library(dplyr)
library(tidyr)
```

# AE Mocks

Here is some mock data created using the `crossing()` function from the `tidyr` package.

```{r mock data}
df <- crossing(label = c("label 1", "label 2", "label 3"),
               column = c("trt1", "trt2", "trt1&trt2", "pl"),
               param = c("count", "percent")) %>%
  # assign numerical order
  mutate(ord1 = rep(seq(1:length(unique(.$label))), each = nrow(.)/length(unique(.$label)) ))

df
```

These mock tables don't contain numerical values as none have been supplied to 
the tfrmt specification. If you would like to add values pass an argument 
to the `value` parameter when applying the `tfrmt()` function.

Note: Layering has been used to avoid repeating code; for more information 
please see the "Reuse and layer tfrmts" article.


```{r}
mock_tbl <- function(tfrmt_obj){
  tfrmt(
    # Specify columns in the data
    label = label,
    column = column, 
    param = param,
    sorting_cols = c(ord1),
    
    # Specify body plan
    body_plan = body_plan(
      frmt_structure(group_val = ".default", label_val = ".default",
                     frmt_combine(
                       "{count} {percent}",
                       count = frmt("XXX"),
                       percent = frmt_when("==100"~ frmt(""),
                                           "==0"~ "",
                                           "TRUE" ~ frmt("(XX.X%)"))))
      ))
}

# Print table
mock_tbl() %>%
  print_mock_gt(df)
```

You can adjust column names and structure using the `col_plan` parameter. It 
accepts selection helpers from the `tidyselect` package.

For example, you can select `everything()` and remove columns that `starts_with("ord")`.

To order and rename columns:

```{r}
# this example is currently not working...

mock_tbl_adjust_cols <- function(){
  mock_tbl() %>%
  tfrmt(
    col_plan = col_plan(
      # -starts_with("ord"),
      # span_structure(
      #   "Treatment",
      #   `T1` = `trt1`,
      #   `T2` = `trt2`,
      #   `T1&T2` = `trt1&trt2`
      #   ),
      # span_structure(
      #   "Placebo",
      #   `PL` = `pl`
      #   )
      
      # -starts_with("ord"),
      # T1 = trt1,
      # T2 = trt2,
      # `T1&T2` = `trt1&trt2`,
      # PL = pl,
      
      -starts_with("ord"),
      span_structure(
        "Treatment",
        T1 = trt1,
        T2 = trt2,
        `T1&T2`= `trt1&trt2`
        ),
      span_structure(
        "Placebo",
        PL = pl
        )
      
      )
  ) 
}
  
mock_tbl_adjust_cols() %>%
  print_mock_gt(df)
```

To add a title, subtitle or footer:

```{r}
mock_tbl_add_titles <- function(){
  mock_tbl_adjust_cols() %>%
    tfrmt(
      # Specify title, subtitle, footer
      title = "Table Name",
      subtitle = "Study ID: GSK12345",
      footer = "A footnote about stuff<br/>Here is a new line with more stuff")
}
  
mock_tbl_add_titles() %>%
  print_mock_gt(df)

```

Mock table with no groups:

```{r}
df <- bind_rows(
  crossing(label = c("label 1", "label 2", "label 3"),
           column = c("T1", "T2", "T1&T2", "PL"),
           param = c("count", "percent")) %>%
    mutate(ord1 = rep(seq(1:length(unique(.$label))), each = nrow(.)/length(unique(.$label)) )),
  crossing(label = c("label 1", "label 2", "label 3"),
           column = c("risk T1-PL", "risk T2-PL", "risk T1&T2-PL"),
           param = c("num", "lower", "upper")) %>%
    mutate(ord1 = rep(seq(1:length(unique(.$label))), each = nrow(.)/length(unique(.$label)) ))
  ) %>% arrange_all()

tfrmt(
  title = "Table Name",
  subtitle = "Study ID: GSK12345",
  footer = "A footnote about stuff<br/>Here is a new line with more stuff",
  # Specify columns in the data
  label = "label",
  param = "param",
  column = "column",
  sorting_cols = vars(ord1),
  # Specify value formatting
  body_plan = body_plan(
    frmt_structure(group_val = ".default", label_val = ".default",
                   frmt_combine(
                     "{count} {percent}",
                     count = frmt("XXX"),
                     percent = frmt_when("==100"~ frmt(""),
                                         "==0"~ "",
                                         "TRUE" ~ frmt("(XX.X%)")))),
    frmt_structure(group_val = ".default", label_val = ".default",
                   frmt_combine(
                     "{num} ({lower}, {upper})",
                     num = frmt("XX.X"),
                     lower = frmt_when("==100"~ frmt(""),
                                       "==0"~ "",
                                       "TRUE" ~ frmt("XX.X%")),
                     upper = frmt_when("==100"~ frmt(""),
                                       "==0"~ "",
                                       "TRUE" ~ frmt("XX.X%"))))
    ),
  # Specify column structure
  col_plan = col_plan(
    T1, T2, `T1&T2`, PL,
    everything(),
    -starts_with("ord")
  )
) %>%
  print_mock_gt(df)
```

Mock table with one level of grouping:

```{r}
df <- bind_rows(
  crossing(group = c("grp1", "grp2"),
           label = c("label 1", "label 2", "label 3"),
           column = c("T1", "T2", "T1&T2", "PL"),
           param = c("count", "percent")) %>%
    mutate(ord1 = rep(seq(1:length(unique(.$label))), each = nrow(.)/length(unique(.$label))),
           ord2 = rep(seq(1,length(unique(.$label))), each = length(unique(.$column))*length(unique(.$param)),
                      times = length(unique(.$group))) ),
  crossing(group = c("grp1", "grp2"),
           label = c("label 1", "label 2", "label 3"),
           column = c("risk T1-PL", "risk T2-PL", "risk T1&T2-PL"),
           param = c("num", "lower", "upper")) %>%
    mutate(ord1 = rep(seq(1:length(unique(.$label))), each = nrow(.)/length(unique(.$label))),
           ord2 = rep(seq(1,length(unique(.$label))), each = length(unique(.$column))*length(unique(.$param)),
                      times = length(unique(.$group))) )
  ) %>% arrange_all()

tfrmt(
  title = "Table Name",
  subtitle = "Study ID: GSK12345",
  footer = "A footnote about stuff<br/>Here is a new line with more stuff",
  # Specify columns in the data
  group = group,
  label = "label",
  param = "param",
  column = "column",
  sorting_cols = vars(ord1, ord2),
  row_grp_plan = row_grp_plan( row_grp_structure(group_val = ".default", element_block(post_space = "   ")) ),
  # Specify value formatting
  body_plan = body_plan(
    frmt_structure(group_val = ".default", label_val = ".default",
                   frmt_combine(
                     "{count} {percent}",
                     count = frmt("XXX"),
                     percent = frmt_when("==100"~ frmt(""),
                                         "==0"~ "",
                                         "TRUE" ~ frmt("(XX.X%)")))),
    frmt_structure(group_val = ".default", label_val = ".default",
                   frmt_combine(
                     "{num} ({lower}, {upper})",
                     num = frmt("XX.X"),
                     lower = frmt_when("==100"~ frmt(""),
                                       "==0"~ "",
                                       "TRUE" ~ frmt("XX.X%")),
                     upper = frmt_when("==100"~ frmt(""),
                                       "==0"~ "",
                                       "TRUE" ~ frmt("XX.X%"))))
    ),
  # Specify column structure
  col_plan = col_plan(
    T1, T2, `T1&T2`, PL, `risk T1-PL`, `risk T2-PL`, `risk T1&T2-PL`,
    everything(),
    -starts_with("ord"))
) %>%
  print_mock_gt(df)

```

Mock table with two levels of grouping: (style 1)

```{r}
df <- bind_rows(
  crossing(grp1 = c("group 1.1", "group 1.2"),
           grp2 = c("group 2.1", "group 2.2"),
           label = c("label 1", "label 2", "label 3"),
           column = c("T1", "T2", "T1&T2", "PL"),
           param = c("count", "percent")),
  crossing(grp1 = c("group 1.1", "group 1.2"),
           grp2 = c("group 2.1", "group 2.2"),
           label = c("label 1", "label 2", "label 3"),
           column = c("risk T1-PL", "risk T2-PL", "risk T1&T2-PL"),
           param = c("num", "lower", "upper"))
  ) %>% arrange_all()

tfrmt(
  title = "Table Name",
  subtitle = "Study ID: GSK12345",
  footer = "A footnote about stuff<br/>Here is a new line with more stuff",
  # Specify columns in the data
  group = c(grp1, grp2),
  label = "label",
  param = "param",
  column = "column",
  row_grp_plan = row_grp_plan( row_grp_structure(group_val = ".default", element_block(post_space = "   ")) ),
  # Specify value formatting
  body_plan = body_plan(
    frmt_structure(group_val = ".default", label_val = ".default",
                   frmt_combine(
                     "{count} {percent}",
                     count = frmt("XXX"),
                     percent = frmt_when("==100"~ frmt(""),
                                         "==0"~ "",
                                         "TRUE" ~ frmt("(XX.X%)")))),
    frmt_structure(group_val = ".default", label_val = ".default",
                   frmt_combine(
                     "{num} ({lower}, {upper})",
                     num = frmt("XX.X"),
                     lower = frmt_when("==100"~ frmt(""),
                                       "==0"~ "",
                                       "TRUE" ~ frmt("XX.X%")),
                     upper = frmt_when("==100"~ frmt(""),
                                       "==0"~ "",
                                       "TRUE" ~ frmt("XX.X%"))))
    ),
  # Specify column structure
  col_plan = col_plan(
    T1, T2, `T1&T2`, PL, `risk T1-PL`, `risk T2-PL`, `risk T1&T2-PL`,
    everything())
) %>%
  print_mock_gt(df)
```

Mock table with two levels of grouping: (style 2)

```{r}
df <- bind_rows(
  crossing(grp1 = c("group 1.1", "group 1.2"),
           grp2 = c("group 2.1", "group 2.2"),
           label = c("label 1", "label 2", "label 3"),
           column = c("T1", "T2", "T1&T2", "PL"),
           param = c("count", "percent")),
  crossing(grp1 = c("group 1.1", "group 1.2"),
           grp2 = c("group 2.1", "group 2.2"),
           label = c("label 1", "label 2", "label 3"),
           column = c("risk T1-PL", "risk T2-PL", "risk T1&T2-PL"),
           param = c("num", "lower", "upper"))
  ) %>% arrange_all()

tfrmt(
  title = "Table Name",
  subtitle = "Study ID: GSK12345",
  footer = "A footnote about stuff<br/>Here is a new line with more stuff",
  # Specify columns in the data
  group = c(grp1, grp2),
  label = "label",
  param = "param",
  column = "column",
  row_grp_plan = row_grp_plan(label_loc = element_row_grp_loc(location = "column")),
  # Specify value formatting
  body_plan = body_plan(
    frmt_structure(group_val = ".default", label_val = ".default",
                   frmt_combine(
                     "{count} {percent}",
                     count = frmt("XXX"),
                     percent = frmt_when("==100"~ frmt(""),
                                         "==0"~ "",
                                         "TRUE" ~ frmt("(XX.X%)")))),
    frmt_structure(group_val = ".default", label_val = ".default",
                   frmt_combine(
                     "{num} ({lower}, {upper})",
                     num = frmt("XX.X"),
                     lower = frmt_when("==100"~ frmt(""),
                                       "==0"~ "",
                                       "TRUE" ~ frmt("XX.X%")),
                     upper = frmt_when("==100"~ frmt(""),
                                       "==0"~ "",
                                       "TRUE" ~ frmt("XX.X%"))))
    ),
  # Specify column structure
  col_plan = col_plan(
    T1, T2, `T1&T2`, PL, `risk T1-PL`, `risk T2-PL`, `risk T1&T2-PL`,
    everything())
) %>%
  print_mock_gt(df)
```



