---
title: "Page Plan"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Page Plan}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, message = FALSE, warning = FALSE}
library(tfrmt)
library(dplyr)
library(gt)
library(tidyr)
library(purrr)
```

The purpose of the page plan is to provide the option to split tables onto multiple pages. Currently, the page plan splits tables horizontally (i.e. row-wise), but in the future it is also planned to allow vertical (i.e. column-wise) splitting. 

## Page Plan rules 

The rules for how to split the table fall into 2 categories:

1) **Values-driven**:

This feature involves splitting by the values of `group` and/or `label` variables. It is specified via one or more `page_structure` objects. Within the `page_structure`, there are two possible methods for defining the splits:

  - Split by every unique value of a `group` or `label` variable. This can be done by passing the value ".default" as such: `page_structure(group_val = ".default")`. 
  
  - Split by a specific value of a `group` or `label` variable. This can be done by passing the specific value as such: `page_structure(label_val = "n")`. This will split after rows where label_val = "n". If there are consecutive rows with label_val = "n", it will split after the final row of that section of rows. This logic extends to instances where there are multiple sections with consecutive rows with label_val = "n"; there will be a split at the bottom of each section. Note: There can only be one `page_structure` that includes this level of specificity.
  
These methods can also be combined with a `page_structure` or across multiple `page_structure`s. For example, `page_structure(group_val = ".default", label_val = "n")` will split on every unique group_val, as well as every instance where label_val = "n". 

If any of the `page_structure`s contain a ".default", it may be desired to print a note indicating the grouping value for the given page at the time of rendering. The location of this note can be specified via the `note_loc` parameter in `page_plan`. The functionality for `note_loc` may be limited by the desired output type; for example, "preheader" is only available for RTF outputs, while "source_note" and "subtitle" are available for all output types. 
  
2) **Max Rows-driven**:

This feature, available via the `max_rows` argument in `page_plan`, involves splitting based on the maximum number of rows per table. Rows dedicated to group labels (without data) are included in the row counts. If a set of rows within a single `group` value are split apart, the group label will be repeated for each page. 

**NOTE**: If both `max_rows` and `page_structure` are provided to the `page_plan`, the table will first be split according to the `page_structure`, follwed by the `max_rows`.


Examples of each of these approaches are below. To reduce the amount of code displayed in the examples, following the initial table, `tfrmt`'s layering functionality will be used to add the `page_plan`.  For more information about this, see the "Layering tfrmts" vignette. 

### Values-driven splitting

Let's take a subset of our example demography data. 
<details>

<summary>Expand for the code used to produce this subset</summary>

```{r}
data_demog2 <- data_demog %>% 
  filter(rowlbl1 %in% unique(rowlbl1)[1:3])
```

</details>

```{r}
head(data_demog2)
```

The formatted table as a single page is as follows:
```{r}
base_tfrmt <- tfrmt(
  # specify columns in the data
  group = c(rowlbl1,grp),
  label = rowlbl2,
  column = column, 
  param = param,
  value = value,
  sorting_cols = c(ord1, ord2),
  # specify value formatting 
  body_plan = body_plan(
    frmt_structure(group_val = ".default", label_val = ".default", frmt_combine("{n} {pct}", 
                                                                                n = frmt("xxx"),
                                                                                pct = frmt_when("==100" ~ "",
                                                                                                "==0" ~ "",
                                                                                                TRUE ~ frmt("(xx.x %)")))),
    frmt_structure(group_val = ".default", label_val = "n", frmt("xxx")),
    frmt_structure(group_val = ".default", label_val = c("Mean", "Median", "Min","Max"), frmt("xxx.x")),
    frmt_structure(group_val = ".default", label_val = "SD", frmt("xxx.xx")),
    frmt_structure(group_val = ".default", label_val = ".default", p = frmt("")),
    frmt_structure(group_val = ".default", label_val = c("n","<65 yrs","<12 months","<25"), p = frmt_when(">0.99" ~ ">0.99",
                                                                                                          "<0.001" ~ "<0.001",
                                                                                                          TRUE ~ frmt("x.xxx", missing = "")))
  ),
  # remove extra cols
  col_plan = col_plan(-grp, 
                      -starts_with("ord") ),
  # Specify column styling plan
  col_style_plan = col_style_plan(
    col_style_structure(align = c(".",","," "), col = vars(Placebo, contains("Dose"), "Total", "p-value"))
  ),
  # Specify row group plan
  row_grp_plan = row_grp_plan(
    row_grp_structure(group_val = ".default", element_block(post_space = " ")),
    label_loc = element_row_grp_loc(location = "column")
  )
)  

base_tfrmt %>% 
  print_to_gt(data_demog2) 
```

#### Every unique value of 1 variable

Suppose we want to split the table by every unique value of `rowlbl1` and add a footnote indicating the table grouping. We can drop `rowlbl1` from the tables since its value will be printed in the note. We will use the `gt::grp_pull()` function to print the individual tables nicely in the vignette. 
```{r  }
gts <- base_tfrmt %>% 
  layer_tfrmt(
    tfrmt(
      # page plan
      page_plan = page_plan(
        page_structure(group_val = list(rowlbl1 = ".default")),
        note_loc = "source_note"
      )   
    )
  ) %>% 
  print_to_gt(data_demog2)
```

```{r}
gts %>% gt::grp_pull(1)
```
```{r}
gts %>% gt::grp_pull(2)
```
```{r}
gts %>% gt::grp_pull(3)
```

#### Every unique value of 2 variables

We could also choose to split on both grouping variables as such (showing first 3 tables only):
```{r}
gts <- base_tfrmt %>% 
  layer_tfrmt(
    tfrmt(
      # page plan
      page_plan = page_plan(
        page_structure(group_val = ".default"),
        note_loc = "source_note"
      )   
    )
  ) %>% 
  print_to_gt(data_demog2) 
```


```{r}
gts %>% gt::grp_pull(1)
```
```{r}
gts %>% gt::grp_pull(2)
```
```{r}
gts %>% gt::grp_pull(3)
```

#### Specific value of a variable

Finally, we could split on a specific value observed in the data. For example, rowlbl1 = "Age (y)".
```{r}
gts <- base_tfrmt %>% 
  layer_tfrmt(
    tfrmt(
      # page plan
      page_plan = page_plan(
        page_structure(group_val = list(rowlbl1 = "Age (y)")),
        note_loc = "source_note"
      )
    )
  ) %>% 
  print_to_gt(data_demog2) 

```


```{r}
gts %>% gt::grp_pull(1)
```
```{r}
gts %>% gt::grp_pull(2)
``` 

### Splitting based on maximum rows

To instead limit the number of rows per page, we can set the `max_rows` argument as such (showing first 3 tables only):

```{r}
gts <- base_tfrmt %>% 
  layer_tfrmt(
    tfrmt(
      # page plan
      page_plan = page_plan(
        max_rows = 20
      )
    )
  ) %>% 
  print_to_gt(data_demog2) 
```

```{r}
gts %>% gt::grp_pull(1)
```
```{r}
gts %>% gt::grp_pull(2)
```  
```{r}
gts %>% gt::grp_pull(3)
``` 


Notice that for groups (defined by the rowlbl1 variable) that are split up, the group header is repeated for each table. 

## Page Plan outputs

When Page Plan is applied, the result of the `print_to_gt` or `print_mock_gt` functions is a `gt_group` object (collection of individual `gt`s). Passing this result to the `gt::gtsave()` function will result in a multi-paged output.
